---
title: "In Class Exercise 4 - Geographically Weighted Summary Statistics - gwModel Methods"
author: "Kieren Chua"
date: "September 20 2024"
date-modified: "last-modified"
execute: 
    eval: true # evaulate the code first
    echo: true # See the code output
    message: false # don't see the warnings
    freeze: true # Prevent re-render
---

# Global Measures of Spatial Autocorrelation

## Part 1 : Initialization
```{r}
# Load the required packages
pacman::p_load(sf, spdep, tmap, tidyverse)
```

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
hunan <- left_join(hunan,hunan2012) %>%
  select(1:4, 7, 15)
```
## Part 2 : Visualising Regional Development Indicator

```{r}
equal <- tm_shape(hunan) +
  tm_fill("GDPPC",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal interval classification")

quantile <- tm_shape(hunan) +
  tm_fill("GDPPC",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal quantile classification")

tmap_arrange(equal, 
             quantile, 
             asp=1, 
             ncol=2)
```

### Computing Contiguity Spatial Weights

#### Queen
```{r}
wm_q <- poly2nb(hunan, 
                queen=TRUE)
summary(wm_q)
```

#### Rook
```{r}
rswm_q <- nb2listw(wm_q, 
                   style="W", 
                   zero.policy = TRUE)
rswm_q
```

### Global Measures of Spatial Autocorrelation: Moran’s I

```{r}
moran.test(hunan$GDPPC, 
           listw=rswm_q, 
           zero.policy = TRUE, 
           na.action=na.omit)
```

**Question** : Based on the value of 0.3, the GDDPC seems to be weakly clustered by location

```{r}
set.seed(1234)
bperm = moran.mc(hunan$GDPPC, 
           listw=rswm_q,
           nsim=999,
           zero.policy = TRUE, 
           na.action=na.omit)
bperm
```

#### Visualization

```{r}
mean(bperm$res[1:999])
var(bperm$res[1:999])
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, 
     freq=TRUE, 
     breaks=20, 
     xlab="Simulated Moran's I")
abline(v=0, 
       col="red") 
```

### Global Measures of Spatial Autocorrelation: Geary’s C

```{r}
geary.test(hunan$GDPPC, listw=rswm_q)
```

Question : Since C value approaches 0, GDPPC seems to be clustered by location

```{r}
set.seed(1234)
bperm=geary.mc(hunan$GDPPC, 
               listw=rswm_q, 
               nsim=999)
bperm
```

#### Visualization

```{r}
mean(bperm$res[1:999])
var(bperm$res[1:999])
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, freq=TRUE, breaks=20, xlab="Simulated Geary c")
abline(v=1, col="red") 
```

## Part 3 : Spatial Correlogram

### Moran I correlogram
```{r}
MI_corr <- sp.correlogram(wm_q, 
                          hunan$GDPPC, 
                          order=6, 
                          method="I", 
                          style="W")
plot(MI_corr)
```

```{r}
print(MI_corr)
```

Question : Since P value < 0.05 for lags of 1, 2, 3, 5, 6, we can determine that for these lags that these points are clustered according to the Moran I statistic and we can reject the NULL hypothesis which is that these points are randomly positioned

### Compute Geary’s C correlogram and plot

```{r}
GC_corr <- sp.correlogram(wm_q, 
                          hunan$GDPPC, 
                          order=6, 
                          method="C", 
                          style="W")
plot(GC_corr)
```

```{r}
print(GC_corr)
```

# Local Measures of Spatial Autocorrelation

## Computing local Moran’s I
```{r}
fips <- order(hunan$County)
localMI <- localmoran(hunan$GDPPC, rswm_q)
head(localMI)
```

```{r}
printCoefmat(data.frame(
  localMI[fips,], 
  row.names=hunan$County[fips]),
  check.names=FALSE)
```

```{r}
hunan.localMI <- cbind(hunan,localMI) %>%
  rename(Pr.Ii = Pr.z....E.Ii..)
```

### Mapping Local Moran Statistics and their P Values

Mapping Local Moran Statistics
```{r}
tm_shape(hunan.localMI) +
  tm_fill(col = "Ii", 
          style = "pretty",
          palette = "RdBu",
          title = "local moran statistics") +
  tm_borders(alpha = 0.5)
```

Mapping their P-values

```{r}
tm_shape(hunan.localMI) +
  tm_fill(col = "Pr.Ii", 
          breaks=c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
          palette="-Blues", 
          title = "local Moran's I p-values") +
  tm_borders(alpha = 0.5)
```

Now together

```{r}
localMI.map <- tm_shape(hunan.localMI) +
  tm_fill(col = "Ii", 
          style = "pretty", 
          title = "local moran statistics") +
  tm_borders(alpha = 0.5)

pvalue.map <- tm_shape(hunan.localMI) +
  tm_fill(col = "Pr.Ii", 
          breaks=c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
          palette="-Blues", 
          title = "local Moran's I p-values") +
  tm_borders(alpha = 0.5)

tmap_arrange(localMI.map, pvalue.map, asp=1, ncol=2)
```

## Creating a LISA Cluster Map

### Plotting Moran scatterplot

```{r}
nci <- moran.plot(hunan$GDPPC, rswm_q,
                  labels=as.character(hunan$County), 
                  xlab="GDPPC 2012", 
                  ylab="Spatially Lag GDPPC 2012")
```

### Plotting Moran scatterplot with standardised variable

```{r}
hunan$Z.GDPPC <- scale(hunan$GDPPC) %>% 
  as.vector 
```

```{r}
nci2 <- moran.plot(hunan$Z.GDPPC, rswm_q,
                   labels=as.character(hunan$County),
                   xlab="z-GDPPC 2012", 
                   ylab="Spatially Lag z-GDPPC 2012")
```

### Preparing LISA map classes

Not sure why to do these steps, will add explanation JIC

```{r}
quadrant <- vector(mode="numeric",length=nrow(localMI))
```

Next, derives the spatially lagged variable of interest (i.e. GDPPC) and centers the spatially lagged variable around its mean.
```{r}
hunan$lag_GDPPC <- lag.listw(rswm_q, hunan$GDPPC)
DV <- hunan$lag_GDPPC - mean(hunan$lag_GDPPC)     
```

This is follow by centering the local Moran’s around the mean.
```{r}
LM_I <- localMI[,1] - mean(localMI[,1])    
```

Set the significance for hypothesis testing
```{r}
signif <- 0.05       
```

These four command lines define the low-low (1), low-high (2), high-low (3) and high-high (4) categories.
```{r}
quadrant[DV <0 & LM_I>0] <- 1
quadrant[DV >0 & LM_I<0] <- 2
quadrant[DV <0 & LM_I<0] <- 3  
quadrant[DV >0 & LM_I>0] <- 4      
```

Lastly, places non-significant Moran in the category 0.
```{r}
quadrant[localMI[,5]>signif] <- 0
```

#### Plot

```{r}
hunan.localMI$quadrant <- quadrant
colors <- c("#ffffff", "#2c7bb6", "#abd9e9", "#fdae61", "#d7191c")
clusters <- c("insignificant", "low-low", "low-high", "high-low", "high-high")

tm_shape(hunan.localMI) +
  tm_fill(col = "quadrant", 
          style = "cat", 
          palette = colors[c(sort(unique(quadrant)))+1], 
          labels = clusters[c(sort(unique(quadrant)))+1],
          popup.vars = c("")) +
  tm_view(set.zoom.limits = c(11,17)) +
  tm_borders(alpha=0.5)
```