---
title: "Take Home Exercise 1"
author: "Kieren Chua"
date: "August 26 2024"
date-modified: "last-modified"
execute: 
    eval: true # evaulate the code first
    echo: true # See the code output
    message: false # don't see the warnings
    freeze: true # Prevent re-render
---

# Take Home Exercise 1

## Part 1 : Reading the Data

```{r}
pacman::p_load("sf", "tidyverse", "tmap", "ggstatsplot")
```

Import the road accident data

```{r}
rdacc_sf <- read_csv("data/thai_road_accident_2019_2022.csv") %>% # Read the data
    filter(!is.na(longitude) & longitude != "", !is.na(latitude) & latitude != "") %>% # Filter for N.A. and empty values
    st_as_sf(coords = c("longitude", "latitude"), crs=4326) %>% # Convert to SF format
    st_transform(crs = 32647) # Convert to the correct format
```

View the data, check for data loss **NO MORE THAN 25%**

```{r}
# Change to dataframe, check col names 
rdacc_df <- as.data.frame(rdacc_sf)
colnames(rdacc_df)
```

```{r}
# Inspect data

# Get descriptions on types of data types
print('presumed_cause')
unique(rdacc_df$presumed_cause)

print('number_of_fatalities')
unique(rdacc_df$number_of_fatalities)

print('weather_condition')
unique(rdacc_df$road_description)

print('weather_condition')
unique(rdacc_df$weather_condition)

print('vehicle_type')
unique(rdacc_df$vehicle_type)
```
May need to bin certain data into smaller bits so as to draw conclusions

```{r}
rdacc_df <- as.data.frame(rdacc_sf)
colnames(rdacc_df)
```

Read the road data

```{r}
thai_rl_sf <- st_read(dsn = "data", layer = "hotosm_tha_roads_lines_shp") %>% st_transform(crs = 32647)

``` 

We are only interested in the Thai Metro Area, so need to read the boundaries from the boundary data and filter there

```{r}
# Read from the province layer, note that the Thai Metro area contains only
    # Bangkok
    # Nonthaburi
    # Nakhon Pathom
    # Pathum Thani
    # Samut Prakan
    # Samut Sakhon
thai_boundaries <- st_read(dsn = "data", layer = "tha_admbnda_adm1_rtsd_20220121") %>% st_transform(crs = 32647)

```
## Part 2 : Initial Plot
We can plot the data so that we can see any kernel clusters

```{r}
# Road accident Data
plot(rdacc_sf)
plot(thai_roadlines)
```




Convert to tibbler dataframe
