---
title: "Hands on Exercise 2, Part 1"
author: "Kieren Chua"
date: "August 28 2024"
date-modified: "last-modified"
execute: 
    eval: true # evaulate the code first
    echo: true # See the code output
    message: false # don't see the warnings
    freeze: true # Prevent re-render
---

# Part 1: *Spatial Point Analysis **
Here we are trying to find out if there are any childcare center cluster and if so, where are they?

```{r}
pacman::p_load(sf, raster, spatstat, tmap, tidyverse)
```

Load in the childcare data
```{r}
childcare_sf <- st_read("data/child-care-services-geojson.geojson") %>% st_transform(crs=3414)
sg_sf <- st_read("data/", layer="CostalOutline")  %>% st_transform(crs=3414)
mpsz_sf <- st_read("data/", layer="MP14_SUBZONE_WEB_PL") %>% st_transform(crs=3414)

colnames(childcare_sf)
colnames(mpsz_sf)
```

## Hands on - Make the point map
```{r}
# Plot the childcare map
tm_shape(mpsz_sf) + tm_polygons() + tm_shape(childcare_sf) + tm_dots(col="black", size=0.05)
```
```{r}
# Pin Map
tmap_mode('view')
tm_shape(childcare_sf)+
  tm_dots()
tmap_mode('plot')
```

# Part 2 : *Geo-spatial Data Wrangling*
```{r}
# Convert to Spatial Data
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```


Display their information
```{r}
childcare
```
```{r}
mpsz
```
```{r}
sg
```
## Convert to SP format
```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

```{r}
childcare_sp
```

```{r}
sg_sp
```
** The difference between Spatial Polygons and Points is the nature of the data, where polygons represent polygons and point is point data. For SpatialPolygonsDataFrame and SpatialPolgons, the one without dataframe just contains the geometric info, while the dataframe contains the data for each polygon **

## Convert to spatstat format

```{r}
childcare_ppp <- as.ppp(childcare_sf)
childcare_ppp
plot(childcare_ppp)
```

```{r}
summary(childcare_ppp)
```
## Checking for duplication and coincidence

```{r}
# Make new dataframe to check
any(duplicated(cbind(childcare_ppp$x, childcare_ppp$y)))
sum(multiplicity(cbind(childcare_ppp$x, childcare_ppp$y)) > 1)
```
** Note : Trick Question, 128 duplicates shown **

```{r}
# Visualize the map
tmap_mode('view')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
tmap_mode('plot')
```

** To find duplicates, find points that are of darker shade than the others, since the alpha made the opacity of the points higher, duplicate points where the points directly overlay each other should be darker **

## Jittering Approach

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
any(duplicated(cbind(childcare_ppp_jit$x, childcare_ppp_jit$y)))
```

## Creating owin object

```{r}
sg_owin <- as.owin(sg_sf)
plot(sg_owin)
```
```{r}
summary(sg_owin)
```
```{r}
childcareSG_ppp = childcare_ppp[sg_owin]
```
```{r}
summary(childcareSG_ppp)
```
Plot the new graph below

```{r}
plot(childcareSG_ppp, frame = FALSE,main="childcareSG_ppp")
plot(sg_owin, add=TRUE)

```
```{r}
summary(childcare_ppp)
```

# Part 3:  First-order Spatial Point Patterns Analysis

```{r}
kde_childcareSG_bw <- density(childcareSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
plot(kde_childcareSG_bw)
```
```{r}
bw <- bw.diggle(childcareSG_ppp)
bw
```
## Rescale KDE values

```{r}
childcareSG_ppp.km <- rescale.ppp(childcareSG_ppp, 1000, "km")
kde_childcareSG.bw <- density(childcareSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG.bw)
```

## Working with different automatic badwidth method

```{r}
# Other methods
bw.CvL(childcareSG_ppp.km) 
bw.scott(childcareSG_ppp.km)
bw.ppl(childcareSG_ppp.km)
bw.diggle(childcareSG_ppp.km)
```

```{r}
# Plotting other methods
kde_childcareSG.ppl <- density(childcareSG_ppp.km, 
                               sigma=bw.ppl, 
                               edge=TRUE,
                               kernel="gaussian")
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "bw.diggle")
plot(kde_childcareSG.ppl, main = "bw.ppl")
```


```{r}
par(mfrow=c(2,2))
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="gaussian"), 
     main="Gaussian")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="epanechnikov"), 
     main="Epanechnikov")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="quartic"), 
     main="Quartic")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="disc"), 
     main="Disc")
```


# Part 4 : Fixed and Adaptive KDE

```{r}
kde_childcareSG_600 <- density(childcareSG_ppp.km, sigma=0.6, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG_600)
```

```{r}
# Adaptive
kde_childcareSG_adaptive <- adaptive.density(childcareSG_ppp.km, method="kernel")
plot(kde_childcareSG_adaptive)
```

## Comparing Fixed and Adaptive

```{r}
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "Fixed bandwidth")
plot(kde_childcareSG_adaptive, main = "Adaptive bandwidth")
```

## Converting KDE Output into GRID

```{r}
gridded_kde_childcareSG_bw <- as(kde_childcareSG.bw, "SpatialGridDataFrame")
spplot(gridded_kde_childcareSG_bw)
```
**TRICK QUESTION**

```{r}
kde_childcareSG_bw_raster <- raster(kde_childcareSG.bw)
kde_childcareSG_bw_raster
```
```{r}
# Assign Projection
projection(kde_childcareSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_childcareSG_bw_raster
```

```{r}
tm_shape(kde_childcareSG_bw_raster) + 
  tm_raster("layer", palette = "viridis") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)
```
```{r}
pg <- mpsz_sf %>%
  filter(PLN_AREA_N == "PUNGGOL")
tm <- mpsz_sf %>%
  filter(PLN_AREA_N == "TAMPINES")
ck <- mpsz_sf %>%
  filter(PLN_AREA_N == "CHOA CHU KANG")
jw <- mpsz_sf %>%
  filter(PLN_AREA_N == "JURONG WEST")
```

```{r}
par(mfrow=c(2,2))
plot(pg, main = "Ponggol")
```
```{r}
plot(tm, main = "Tampines")
```
```{r}
plot(ck, main = "Choa Chu Kang")
```

```{r}
plot(jw, main = "Jurong West")
```

```{r}
# Create OWIN object to perform KDE on the idividual locations
pg_owin = as.owin(pg)
tm_owin = as.owin(tm)
ck_owin = as.owin(ck)
jw_owin = as.owin(jw)
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]
childcare_pg_ppp.km = rescale.ppp(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale.ppp(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale.ppp(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale.ppp(childcare_jw_ppp, 1000, "km")
par(mfrow=c(2,2))
plot(childcare_pg_ppp.km, main="Punggol")
plot(childcare_tm_ppp.km, main="Tampines")
plot(childcare_ck_ppp.km, main="Choa Chu Kang")
plot(childcare_jw_ppp.km, main="Jurong West")
```
```{r}
par(mfrow=c(2,2))
plot(density(childcare_pg_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tempines")
plot(density(childcare_ck_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Choa Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Jurong West")
```
```{r}
# Use 250m as bandwidth
par(mfrow=c(2,2))
plot(density(childcare_ck_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Chou Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
plot(density(childcare_pg_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tampines")
```

# Part 5 : Nearest Neighbour Analysis

```{r}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```
**Can see that there is a clustering of childcare centers in the total of sg**


```{r}
clarkevans.test(childcare_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

```{r}
clarkevans.test(childcare_tm_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```