---
title: "Take Home Exercise 2"
author: "Kieren Chua"
date: "August 26 2024"
date-modified: "September 28 2024"
execute: 
    eval: true # evaulate the code first
    echo: true # See the code output
    message: false # don't see the warnings
    freeze: true # Prevent re-render
---

# Part 1 : Reading the Data

```{r}
pacman::p_load("sf", "tidyverse", "tmap", "ggstatsplot", "spatstat", "tmaptools")
```

```{r}
# Read the provincial data and the other data
thailand_csv <- read_csv("data/thailand_domestic_tourism_2019_2023_ver2.csv") %>% mutate(year = year(date)) %>% mutate(month = month(date, label=TRUE)) %>% select(-province_thai, -region_thai)

province_shp <- st_read(dsn = "data", layer = "tha_admbnda_adm1_rtsd_20220121") %>% st_transform(crs = 32647)
```
```{r}
# See number of months and number of years

```

```{r}
# Combine the data
# combined_data <- left_join(province_shp, thailand_csv, by=("ADM1_EN" = thailand_csv$province_eng))
```

# Part 2 : Check if Space & Time - Independent

To check if the data is space independent, we need to see if there is any correlation between the tourist numbers of each region to its neighbors. We shall use a non-COVID time period which is the year of 2019, so that the data is not affected by external factors.

```{r}
thailand_csv_2019 <- thailand_csv %>% filter(year == 2019)
```

After splitting by month, we can plot a space graph by month and then plot the number of tourist 
by the month and location simultaneously. If we see any variation of clusters by the month, we know that it is time
dependent. If there is a consistent cluster among all the time location, we know that it is space dependent.

## Determining Neighbours

```{r}
```



