---
title: "Take Home Exercise 2"
author: "Kieren Chua"
date: "August 26 2024"
date-modified: "September 28 2024"
execute: 
    eval: true # evaulate the code first
    echo: true # See the code output
    message: false # don't see the warnings
    freeze: true # Prevent re-render
---

# Part 1 : Reading the Data

```{r}
pacman::p_load("sf", "sfdep", "tidyverse", "tmap", "ggstatsplot", "spatstat", "tmaptools")
```

```{r}
# Read the provincial data and the other data
thailand_csv <- read_csv("data/thailand_domestic_tourism_2019_2023_ver2.csv") %>% 
                mutate(year = year(date)) %>% 
                mutate(month = month(date, label=TRUE))

# There is some missing data
unqiue_province <- unique(thailand_csv$province_eng)

province_shp <- st_read(dsn = "data", layer = "tha_admbnda_adm1_rtsd_20220121") %>% 
                st_transform(crs = 32647) %>%
                select(-ADM1_REF, 
                       -ADM1ALT1EN, 
                       -ADM1ALT2EN, 
                       -ADM1ALT1TH, 
                       -ADM1ALT2TH, 
                       -ADM0_TH, 
                       -ADM0_EN,
                       -ADM0_PCODE,
                       -ADM1_TH) %>%
                rename(province_eng = ADM1_EN) %>% # Rename for spacetime
                filter(province_eng %in% unqiue_province) %>% #Remove missing
                mutate(centroid = st_centroid(geometry))

# Check for empty geometries
empty_geometries <- st_is_empty(province_shp)
if (any(empty_geometries)) {
    warning("Some geometries are empty.")
} else {
  print("No Empty Geometries")
}
```
Now we can draw out the data 
```{r}
tmap_mode("plot")
tm_shape(province_shp) +
  tm_polygons()
```
```{r}
tmap_mode("plot")
```

# Part 2 : Determining Key-Economic Indicators

```{r}
economic_inidcators <- unique(thailand_csv$variable)
print(economic_inidcators)
```

Looking at the indicators


# Part 2 : Check if Space - Independent

We can check if the location is space dependent by fixing the location to a specific time zone which would represent the general trend without the effect of time. This year would be 2019 when CoVID was not prevelant

# Part 3 : Check if Space & Time - Independent

To check if the data is space independent, we need to see if there is any correlation between the tourist numbers of each region to its neighbors. We can split it between 3 time period, which is Pre-Covid (2019), Covid (2020 - 2022) and Post - Covid (2023) to observe the effect of time without the influence of CoVID.

## Creating the Space-Time Layer

```{r}
spacetime_thai_year <- spacetime(thailand_csv, 
                                    .loc_col = "province_eng", 
                                    .time_col= "year",
                                    .geometry = province_shp)
```

Verify that spacetime object has both geometry and data

```{r}
activate(spacetime_thai_year, "data")
```

```{r}
activate(spacetime_thai_year, "geometry")
```

## Global Moran's I Test

```{r}
# TODO : Change to Distance Metrics, Use KnearestNeighbours, and change to centroids
wm_q <- province_shp %>% 
        mutate(nb = st_contiguity(geometry),
                wt = st_weights(nb, style = 'W'), # Standardised over neighbours
                .before = 1,
                #zero.viewpolicy = TRUE,
                allow_zero = TRUE) # There seems to be some islands
```

Now we conduct Moran's I for every month and every year
```{r}
unique_month <- unique(thailand_csv$month)
unique_year <- unique(thailand_csv$year) 
```

```{r}
```
